paper

### 2019-01-11 Report/paper.tex
push paper and figures

\begin{abstract}
    In this report we explore ways to automatically analyze ICLS and CSCL
papers to gain an understanding of the community formed by conference
participants and of the research being conducted within the community.
After describing how to extract data from the base dataset we derive
insights into the community using network analysis, natural language
processing and text mining. We discover patterns of collaboration
between participants, their migration and the global spread of the
community. We extract keywords and find clusters of referenced papers
and documents within the conferences.
\end{abstract}
    \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\begin{itemize}
\tightlist
\item
  The institutions and countries dominant in the conferences
\item
  The differences between ICLS and CSCL with respect to citations,
  research conducted and contributors
\item
  Collaboration network patterns with respect to countries, institutions
  and authors
\item
  First insight into changes over the years with respect to contributors
  and collaboration patterns
\item
  The migration of participants across institutions
\item
  Popularity of keywords over the years
\item
  Trends inregards to citated papers and the evolution of their
  prevalence from 1977 until 2018
\item
  Discovering clusters of the cited documents
\item
  Discovering clusters of the papers written within the conferences
\end{itemize}

@@

    The provided raw dataset consists of papers in \emph{pdf} format, as
well as associated metadata in \emph{xml} format. To extract information
from the dataset we first converted the data to a format that is more
adapt to the task of data analysis and more compatible with a variety of
external libraries. We stored all the resulting data in \emph{csv}
format, as csv is a human readable format and can easily be processed by
all the data analysis tools we employ.

@@

We extracted the metadata from the xml files using the \emph{lxmltree}
python library. We associated the contents of each xml tag to a column
in a csv file. As the xml files do not exactly follow the xml standard,
we handled errors by iteratively parsing the xml tree. For some files,
the section containing the keywords is malformed and can not be
extracted by the parser. This was remedied using regular expressions to
match the section containing the keywords. To each file we associated
the name of the source file and the order in which an author name occurs
in the xml file. Each author is associated to one line in the xml file.
From the associated citation string we add to the first part of the
citation containing the author names, year and title. This is later used
to identify the paper when it is referenced in an other paper. Then we
add a shortened version of author names (i.e.~West, R.) by considering
the author order. Note that it happens that the short name extracted
from the citation and the corresponding long name (Robert West) in the
metadata do not match up. Thus the author order listed in the citation
string does not always correspond to the author order listed in the
metadata. After extraction, we cleaned the data by unifying names, as
slight spelling differences exist within the same author name
(i.e.~Robert West, Robert A. West). When unifying the names we needed to
make sure not to overmerge by labeling two different authors as having
the same name. First, we used a strict condition to catch small
differences in names, such as an extra middle name. This is done by
splitting up names into substrings based on whitespaces and commas. For
two names in the dataset we checked whether the intersection of their
sets of substrings contained at least 2 names of length at least 3. If
the condition was met we considered these two names to be identical, and
added the mapping from one name to the other to a dictionary. Once all
tuples of names (s.t. the first name appears before the second one in
order) were checked and added to the dictionary we unified the names.
Using this method we reduced the number of authors in the dataset from
1951 down to 1879. Further, to merge names with misspellings or people
who use nicknames (Christian vs Chris) and to avoid overmerging we
considered only names of authors that have similar collaboration
patterns. To do this we built a graph based on co-authors relations and
checked names of authors that were in the same neighborhood of a node on
the graph (see methodology for further details). Then, we used the
\emph{difflib} library to check for similarity of two strings and
considered a name to belong to the same person if the similarity measure
reached the threshold of 0.8. Again names considered to belong to the
same person were added to the mapping dictionary which was then applied
to all names. This method again reduced the number of authors in the
dataset from 1879 to 1951.

@@

To extract information from the papers themselves we needed first to
extract the text from the pdfs. Various parsers performing this task
exist. However, parsers such as pdfminer or PyPDF2 that claim to be able
to extract the text as well as metadata from papers have not been kept
up to date, which leads to issues with dependencies. We found that the
most reliable parser is \emph{poppler} \cite{poppler} which can be
installed and then used via the command line to extract the text
contained in the pdfs to \emph{txt} format. All methods describing how
to extract data from the papers worked on the extracted \emph{txt}
files. The characters contained in the \emph{txt} files were not all
encoded uniformly, which means that characters that look the same to
humans may be encoded in different ways. Hence, for many converting all
characters to `Normal Form Composed' form.

@@

\textbf{Regular expressions in python:} Python has two modules that
implement regular expressions, \emph{re} and \emph{regex}. We used regex
as --unlike the re module-- this module allows to handle strings with
non-ascii characters by having unicode expressions. As many authors have
names containing non-ascii characters this is an important capability.

\textbf{To get the reference section for of a paper} we used a regular
expressions to find the reference header and determine the sections end
by checking for acknowledgements or appendix sections in the text file
previously extracted using poppler. Then, to extract individual
references we split the section into a list of substrings at each new
line character and concatenated substrings back together if they
satisfied a set of conditions. We found this method to be robust as it
makes sure that all properly formatted references are separated.

\textbf{To unify substrings} we exploited the fact that APA references
all have the same initial structure: \emph{Lastname F.M. (when
published) Title which may contain colons and other non-alphabetic
chars. Where published. Who published} \cite{apaformat}. We thus checked
that the beginning of each reference contains ``Lastname F.M. (when
published)''. It is relatively easy to check for this using regular
expressions as author names and year have a given structure. If a string
does not contain such a section, then it must be part of an other
reference, which lies before in the list of reference substrings. Note
that this beginning is also what is added to the metadata as identifying
string. We found that in practical application it is important to first
merge lines if they were clearly cut off, as papers with many authors
may have the beginning on multiple lines. We could find split up lines
by checking for lines starting with non-alphabetic characters or
starting in a sentence. We merged these with the line above. We could
identify lines by their ending as being split up, such as if they end in
a comma, colon or only contain author names. We merged these lines with
the lines below them. Only then could we check whether a line contained
a valid citation start. If it did not we merged it with the line above
it. After this merging process we validated the returned reference split
by looking at the overall length and the initials, as well as by
checking that a random sample returned satisfactory splitting accuracy.
We found that the above method of extracting the references produced
good results for proper APA citations and on APA citations indeed
outperformed Scolarcy when it came to splitting up the lines.

@@

The most robust way to get a persons institution affiliation is to use
their email address. Email addresses can easily be extracted from the
header section using a regular expression. Then we used a
domain-university mapping \cite{mapping} to map domains to institutions.
The additional advantage of this method is that using the additional
metadata associated to each university in the mapping, we also access
information about the country of the university.

@@

The header section itself contains information on the institution, but
different ways to list authors are accepted, notably two ways as can be
seen in Figure \ref{figA} and Figure \ref{figB}. Hence, we first
detected the structure of the header section before extracting the
association. We tagged the substring present in this section using a
combination of \emph{named entity detection}, using the implementation
provided by \emph{spacy}, simple string comparison matching of the names
of authors and institutions we extracted from the metadata respectively
from Approach 1 and tagging sections containing ``emails'' using regex ,
to detect where in the header the email is. Then, taking the tagged
section, we detected which subsections corresponded to which
author-affiliation listing type, and extracted the author email -
institution mapping accordingly.

Approach 2 allows to match many more institutions to authors than
Approach 1 as not every institution is present in the mapping --no
companies will be in the mapping-- and many participants use gmail and
similar providers as their official email. However, Approach 2 may
contain differences in spelling in the institution name, which needed to
be merged. Thus we combined both approaches, and used the corresponding
institutions in Approach 1 to unify the institutions obtained in
Approach 2. Furthermore, we later merged names by considering the
location of the institutions.

But first as with Approach 2 institution affiliation did not allow us to
automatically associate a country to the institution, we needed to get
this through other means. A first step was to use the top level domains
present in the emails to associate a country of origin to the
participants and their associated institutions. Then, to plot the
community we augmented the dataset with geolocation information. As a
further benefit, having this information we could then easily associate
a country to the institution for the emails that do not have a top level
domain indicating location and further merge the names of institutions
based on location.

As there are a variety of way in which the name of an institution could
be written --consider EPFL, EPF Lausanne-- crawling the english, french
and german version of Wikipedia (not using the API, or the API wrapper,
as this does not give access to the desired information) was a quick and
accurate way to get the location. The advantage of sending requests over
using the API is due to the website redirecting to the relevant article
on the institution even when misspellings are present. Requests were
sent using \emph{urllib3} and \emph{beautifulsoup} was used to parse
them get the coordinates present in the article. Regretfully not all
institutions, even well known ones such as Tokyo University, have their
coordinates listed on Wikipedia. To get the remaining locations we used
the geopy, which is a client for geocoding web services. These results
however are more error prone and will always return a result. It is thus
recommended to verify the correctness of the results. To get the country
we then again used geopy. The resulting data containing paper
identifier, the authors email, its associated domain, order in which the
author was cited, the institution, country, longitude and latitude were
saved to a csv file. We used the paper identifier and author order to
associate this data with the metadata to link information about authors
with the papers they authored.

    \hypertarget{keywords-extraction}{%
\subsection{Keywords extraction}\label{keywords-extraction}}

@@

    From the pdf parser some of the documents were corrupted with some weird
characters such as , white space strings

\begin{itemize}
\item
  We used \textbf{\emph{beautiful soup}} and \textbf{\emph{string}}
  libraries to only keep the text that we needed without any special
  character.
\item
  we applied \textbf{\emph{regular expression}} to remove numbers
  because we thought that numbers were not that important.
\item
  We put all the words to lower cases
\item
  We removed the words that were connecting parts of a sentence rather
  than showing subjects, objects or intent (\textbf{\emph{stopwords}}).
  For us these words were not that important and were only going to
  increase the bag of words length
\item
  In some methods we use \textbf{\emph{Lemmatization}} and
  \textbf{\emph{stemming}} and in other methods we don't because we
  wanted to do a mapping with \textbf{\emph{GloVe}} vocabulary which we
  will explain later on.
\end{itemize}

\textbf{\emph{Lemmatization}} is a word variations to the root of the
word (e.g.~working, works, worked changed to work).
\textbf{\emph{Stemming}} is the process of reducing inflected and
derived words to their word stem, base or root form---generally a
written word form.

As the use of the stop words to reduce the BOW (bag of words size) we
also used lemmatization and stemming to be able to reduce the number of
words without taking off any valuable information.

    \textbf{\emph{Lemmatization}} and different form of
\textbf{\emph{Stemming}} were only used with \textbf{\emph{TF-IDF}}
while doing the Keywords extraction, otherwise we prefered keeping the
words as they were because we were doing a mapping with a pretrained
stanford word embedding and sometimes root forms were not in the
dictionnary of words of the pre-trained words-vectors.

    For the Keywords extraction we used three methods :

The first one is a trivial one is to search in the pdf itself for
keywords and extract them, this method is the most accurate but we found
out that for 492 papers we had no keywords included. So we deciced to
use two methods for keywords extraction. The first one is
\textbf{\emph{TF-IDF}} and the second one is \textbf{\emph{Rake}}.

    \begin{itemize}
\tightlist
\item
  \textbf{\emph{TF-IDF}}: Tf--idf is one of the most popular
  term-weighting schemes today, in information retrieval, tf--idf or
  TFIDF, short for term frequency--inverse document frequency, is a
  numerical statistic that is intended to reflect how important a word
  is to a document in a collection or corpus. It is often used as a
  weighting factor in searches of information retrieval, text mining,
  and user modeling. The tf--idf value increases proportionally to the
  number of times a word appears in the document and is offset by the
  number of documents in the corpus that contain the word, which helps
  to adjust for the fact that some words appear more frequently in
  general \cite{Tf-idf}.
\end{itemize}

    With TF-IDF we sorted for each document the words with the highest score
and then selected the the 7 first words of each document.

    \begin{itemize}
\tightlist
\item
  \textbf{\emph{Rake}}: Rake stands for Rapid Automatic Keyword
  Extraction. It is an existing python implementation that uses the NLTK
  toolkit for the score calculations.
\end{itemize}

\textbf{\emph{How it works}} :

@@

    Once we have all the keywords from the three methods, we try to select
the most accurate keywords. So we created a method that selects the
intersection of TF-IDF keywords and the Rake keywords with the keywords
of the authors (if there are any keywords in the text), if the
intersection is empty and we have the keywords in the text we only take
the keywords of the text. If the paper does not have keywords, we take
the intersection of the TF-IDF and the Rake. If again, the intersection
is empty then we only take the TF-IDF keywords because they were closer
to the text keywords than the Rake Keywords.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_27_0.png}\end{center}
        \caption{Trends of keywords over the years}
        \label{fig1}
    \end{figure}

    We wanted then to know what are the trends of words for each specific
year, so we select distinct words for each year. On other words it means
that we select for each year the words that were only present for this
specific year and never appeared in the others. We get the figure below.
As we can see from the graph \ref{fig2}, the most used word in the
papers, more than 10\% of 2015 papers uses the word `peer', more than
6\% of the papers uses the words `building', `embodied' in 2016. In 2017
the most frequent keyword on is `reponsive' and in 2018 `tools' is
present with more than 6\%.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_29_0.png}\end{center}
        \caption{Distinct keywords over the years}
        \label{fig2}
    \end{figure}

@@

To cluster the documents we decided to use two different approaches. The
first one is Natural Langage Processing (NLP) and the second one is a
network analysis. The most accurate one is the network clustering using
the references graph.

@@

\hypertarget{baseline-tf-idf-combined-with-k-means}{%
\paragraph{Baseline : TF-IDF combined with
K-means}\label{baseline-tf-idf-combined-with-k-means}}

With the natural langage processing, we started with a naive method
which consists of clustering the documents using only TF-IDF. Each
document has a vector representation with values for each word present
in the document depending on the frequency of the word in this document
regarding its frequency in all the other documents. So if a word is a
keyword to the document and is not a common word than its value will be
more important than the other words in the vector of the document. Once
we obtained our vector representation, we selected the max features
parameter as 2500, it means that we only considered the top
max\_features ordered by term frequency across the corpus. On this word
representation we run K-means.

\textbf{\emph{K-means}} clustering is a method used for clustering
analysis, especially in data mining and statistics. It aims to partition
a set of observations into a number of clusters (k), resulting in the
partitioning of the data into Voronoi cells. It can be considered a
method of finding out which group a certain object really belongs to.

This method did not give good results as we can see it in the graph
below \ref{fig3}. All the vectors are together there is no clear
separation between the documents. So we decide to use other methods
using words embeddings and properties of TF-IDF.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_32_0.png}\end{center}
        \caption{Baseline model document clusters}
        \label{fig3}
    \end{figure}

@@

The second method was to use pre-trained words embeddings. We decided to
use \textbf{\emph{GloVe}} library pre-trained on Wikipedia dataset.

@@

We constructed a dictionnary from the words embeddings mapping each word
of a document to its vector representation, than we averaged all the
vectors of each document. Since each words has its representation in
space, the average of all the words for each document will lead to a
vector representing a document.

Once we obtained a document vector representation, we applied to it
K-means. The results were slightly better, but we had only two clear
separate clusters. If we increase the number of clusters than there is
no clear separation anymore. The plot below \ref{fig4} reprensents the
documents belonging to each cluster.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_34_0.png}\end{center}
        \caption{Cluster using Average of GloVe vectors combined with K-means }
        \label{fig4}
    \end{figure}

@@

With this method, we did not use the parameter max features of the
TF-IDF vector, we took all the vocabulary generated by all the documents
and we created a method that sorts the document frequency vector. We
selected the 500 first words and did a weighted average of the words
embedding vectors with the TF-IDF score.

Once we get the weighted average vector, and since we were using a words
embedding of 50 dimentions. We decided to use feature reduction
technique to reduce the number of dimensions and represent our vectors
in an other vector space where all the features are orthogonal to each
others.

To apply the feature reduction to our vectors we used
\textbf{\emph{Principal Component Analysis (PCA)}} algorithm.

@@

Once we get our new vector representation, we applied K-means and we
obtained the best separation, with a clear separation of 4 clusters as
we can see it in the figure below \ref{fig5}.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_36_0.png}\end{center}
        \caption{Cluster using Weighted average of TF-IDF + GloVe  combined with K-means }
        \label{fig}
    \end{figure}

@@

For the references graph we chose to construct 4 different graphs to be
able to extract the more information we can. Our first purpose is to be
able to cluster the documents that are about the same subject and to
gain more insights about the most influencial papers, trends of
citations\ldots{}

@@

Our first graph is constructed using the references of each document. It
is constructed as follows, for each two documents, if there is an
intersection between the references of both documents then there is a
link. The weights of the edges are the length of the intersection set.
The intuition behind it, is that the more the weight is high, the more
the documents are likely to be about the same subject since they are
citing the same documents. From this graph, we create a subgraph of it,
we filter the edges keeping only those who have a weight strictly higher
than 3. It means that, each node of the subgraph, is related to at least
one other node with 4 same documents referenced.

    The resulted graph is plotted in this figure \ref{intersection}, is an
undirected graph with 805 nodes and 6699 edges. On this graph we extract
two main information : * Strongly connected components to be able to
detect clusters and related documents. * In\_degree in order to check if
there are documents that are related to a lot of documents and those
documents would be more general than others, and won't necessarly belong
to one specific cluster.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_42_0.png}\end{center}
        \caption{Intersection between references graph}
        \label{intersection}
    \end{figure}

@@

>>>>

\begin{abstract}
    In this report we investigate ways to automatically analyze ICLS and
CSCL papers to gain an understanding of the community formed by
conference participants and of the research being conducted within the
community. After describing how to extract data from the base dataset we
derive insights into the community using network analysis, natural
language processing and text mining. We discover patterns of
collaboration between participants, their migration and the global
spread of the community. We extract keywords and find clusters of
referenced papers and documents within the conferences.
\end{abstract}
    \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\begin{itemize}
\tightlist
\item
  The institutions and countries dominant in the conferences
\item
  The differences between ICLS and CSCL with respect to citations,
  research conducted and contributors
\item
  Collaboration network patterns with respect to countries, institutions
  and authors
\item
  First insight into changes over the years with respect to contributors
  and collaboration patterns
\item
  The migration of participants across institutions
\item
  Popularity of keywords over the years
\item
  Trends in regards to cited papers and the evolution of their
  prevalence from 1977 until 2018
\item
  Clusters of the cited documents
\item
  Clusters of the papers written within the conferences
\end{itemize}

@@

    The provided raw dataset consists of papers in \emph{pdf} format, as
well as associated metadata in \emph{xml} format. To extract information
from the dataset we first converted the data to a format that is better
adapted to the task of data analysis and more compatible with a variety
of external libraries. We stored all the resulting data in \emph{csv}
format, as csv is a human readable format and can easily be processed by
all the data analysis tools we employ.

@@

We extracted the metadata from the xml files using the \emph{lxmltree}
python library. We associated the contents of each xml tag to a column
in a csv file. As the xml files do not exactly follow the xml standard,
we handled errors by iteratively parsing the xml tree. For some files,
the section containing the keywords is malformed and can not be
extracted by the parser. This was remedied using regular expressions to
match the section containing the keywords. To each file we associated
the name of the source file and the order in which an author name occurs
in the xml file. Each author is associated to one line in the xml file.
From the associated citation string we add the first part of the
citation containing the author names, year, and title. This is later
used to identify the paper when it is referenced in an other paper. Then
we add a shortened version of author names (i.e.~West, R.) by
considering the author order. Note that it happens that the short name
extracted from the citation and the corresponding long name (Robert
West) in the metadata do not match up. Thus the author order listed in
the citation string does not always correspond to the author order
listed in the metadata. After extraction, we cleaned the data by
unifying names, as slight spelling differences exist within the same
author name (i.e.~Robert West, Robert A. West). When unifying the names
we needed to make sure not to overmerge by labeling two different
authors as having the same name. First, we used a strict condition to
catch small differences in names, such as an extra middle name. This is
done by splitting up names into substrings based on whitespaces and
commas. For two names in the dataset we checked whether the intersection
of their sets of substrings contained at least 2 names of length at
least 3. If the condition was met we considered these two names to be
identical, and added the mapping from one name to the other to a
dictionary. Once all tuples of names (s.t. the first name appears before
the second one in order) were checked and added to the dictionary we
unified the names. Using this method we reduced the number of authors in
the dataset from 1951 down to 1879. Further, to merge names with
misspellings or people who use nicknames (Christian vs Chris) and to
avoid overmerging we considered only names of authors that have similar
collaboration patterns. To do this we built a graph based on
co-authorship relations and checked names of authors that were in the
same neighborhood of a node on the graph (see methodology for further
details). Then, we used the \emph{difflib} library to check the
similarity of two strings and considered a name to belong to the same
person if the similarity measure reached the threshold of 0.8. Again,
names considered to belong to the same person were added to the mapping
dictionary which was then applied to all names. This method again
reduced the number of authors in the dataset from 1879 to 1851.

@@

To extract information from the papers themselves we needed first to
extract the text from the pdfs. Various parsers performing this task
exist. However, parsers such as pdfminer or PyPDF2 that claim to be able
to extract the text as well as metadata from papers have not been kept
up to date, which leads to issues with dependencies. We found that the
most reliable parser is \emph{poppler} \cite{poppler} which can be
installed and then used via the command line to extract the text
contained in the pdfs to \emph{txt} format. All methods describing how
to extract data from the papers worked on the extracted txt files. The
characters contained in the txt files were not all encoded uniformly,
which means that characters that look the same to humans may be encoded
in different ways. Hence, for many applications we converted all
characters to `Normal Form Composed' form.

@@

\textbf{Regular expressions in python:} Python has two modules that
implement regular expressions, \emph{re} and \emph{regex}. We used regex
as --unlike the re module-- this module allows to handle strings with
non-ascii characters by allowing the use of unicode regular expressions.
As many authors have names containing non-ascii characters this is an
important capability.

\textbf{To get the reference section for of a paper} we used a regular
expressions to find the reference header and determine the end of the
section by checking for acknowledgements or appendix sections in the
text file previously extracted using poppler. Then, to extract
individual references we split the section into a list of substrings at
each new line character and concatenated substrings back together if
they satisfied a set of conditions. We found this method to be robust as
it makes sure that all properly formatted references are separated.

\textbf{To unify substrings} we exploited the fact that APA references
all have the same underlying structure: \emph{Lastname F.M. (when
published) Title which may contain colons and other non-alphabetic
chars. Where published. Who published} \cite{apaformat}. We thus checked
that the beginning of each reference contains ``Lastname F.M. (when
published)''. It is relatively easy to check for this using regular
expressions as author names and year have a given structure. If a string
does not contain such a section, then it must be part of an other
reference, which lies before in the list of reference substrings. Note
that this beginning is also what is added to the metadata as identifying
string. We found that in practical application it is important to first
merge lines if they were clearly cut off, as papers with many authors
may have the beginning on multiple lines. We could find split up lines
by checking for lines starting with non-alphabetic characters or
starting in a sentence. We merged these with the line above. We could
identify lines by their ending as being split up, such as if they end in
a comma, colon or only contain author names. We merged these lines with
the lines below them. Only then could we check whether a line contained
a valid citation start. If it did not, we merged it with the line above
it. After this merging process we validated the returned reference split
by looking at the overall length and the initials, as well as by
checking that a random sample returned satisfactory splitting accuracy.
We found that the above method of extracting the references produced
good results for proper APA citations and on APA citations indeed
outperformed Scolarcy when it came to splitting up the lines.

@@

The most robust way to get a persons institution affiliation is to use
their email address. Email addresses can easily be extracted from the
header section using a regular expression. Then, the part of the email
containing the university domain can be extracted using an other regular
expression. After this we used a domain-university mapping
\cite{mapping} to map domains to institutions. The additional advantage
of this method is that using the additional metadata associated to each
university in the mapping, we also access information about the country
of the university.

@@

The header section itself contains information on the institution, but
different ways to list authors are accepted, notably two ways as can be
seen in Figure \ref{figA} and Figure \ref{figB}. Hence, we first
detected the structure of the header section before extracting the
association. We tagged the substring present in this section using a
combination of \emph{named entity detection}, which automatically
classifies strings as belonging to some pre-defined category, using the
implementation provided by \emph{spacy}\cite{spacy2}, and simple string
comparison matching of the names of authors and institutions we
extracted from the metadata respectively from Approach 1 and tagging
sections containing ``emails'' using regex , to detect where in the
header the email is. Then, taking the tagged section, we detected which
subsections corresponded to which author-affiliation listing type, and
extracted the author email - institution mapping accordingly.

Approach 2 allows to match many more institutions to authors than
Approach 1 as not every institution is present in the mapping and many
participants use gmail and similar providers as their official email.
However, Approach 2 may contain differences in spelling in the
institution name, which needed to be unified. Thus we combined both
approaches, and used the corresponding institutions in Approach 1 to
unify the institutions obtained in Approach 2. Furthermore, we later
merged names by considering the location of the institutions.

But first, as with Approach 2 institution affiliation did not allow us
to automatically associate a country to the institution, we needed to
get this information through other means. A first step was to use the
top level domains present in the emails to associate a country of origin
to the participants and their associated institutions. Then, to plot the
community we augmented the dataset with geolocation information. As a
further benefit, having this information we could then easily associate
a country to the institution for the emails that ido not have a top
level domain indicating location and further merge the names of
institutions based on location.

As there are a variety of way in which the name of an institution could
be written --consider EPFL, EPF Lausanne-- crawling the english, french
and german version of Wikipedia (not using the API, or the API wrapper,
as this does not give access to the desired information) was a quick and
accurate way to get the location. The advantage of sending requests over
using the API is due to the website redirecting to the relevant article
on the institution even when misspellings are present or multiple ways
to spell the university exist. Requests were sent using \emph{urllib3},
and \emph{beautifulsoup} was used to parse them get the coordinates
present in the article. Regretfully not all institutions, even well
known ones such as Tokyo University, have their coordinates listed on
Wikipedia. To get the remaining locations we used geopy, which is a
python client for geocoding web services. These results however are more
error prone as geopy will always return a result. It is thus recommended
to verify the correctness of the results. To get the country we then
again used geopy. The resulting data containing paper identifier, the
authors email, its associated domain, order in which the author was
cited, the institution, country, longitude and latitude were saved to a
csv file. We used the paper identifier and author order to associate
this data with the metadata to link information about authors with the
papers they authored.

    \hypertarget{keyword-extraction}{%
\subsection{Keyword extraction}\label{keyword-extraction}}

@@

    The pdf parser returns txt files that are encoded using a variety of
conventions, thus when extracting the strings from the text files, we
were met with non-ascii characters, which had to be dealt with. We
preprocessed the text by applying the following steps to it:

\begin{itemize}
\item
  We used \emph{beautiful soup} and \emph{string} libraries to only keep
  sections of the text that we needed and remove special characters.
\item
  We applied \textbf{\emph{regular expression}} to remove numbers
  because numbers were not important for our analysis.
\item
  We put all words to lower case
\item
  We removed words that were connecting parts of a sentence rather than
  showing subjects, objects or intent (\textbf{\emph{stopwords}}). For
  the purpose of this task these words were not important and were only
  going to increase the bag of words length
\item
  In some methods we applied \textbf{\emph{lemmatization}} and
  \textbf{\emph{stemming}} to the words. In other methods we did not
  because we implement a mapping with \textbf{\emph{GloVe}} vocabulary
  instead, which we will explain later on.
\end{itemize}

\textbf{\emph{Lemmatization}} is a word variations to the root of the
word (e.g.~working, works, worked changed to work).
\textbf{\emph{Stemming}} is the process of reducing inflected and
derived words to their word stem, base or root form---generally a
written word form. As the use of the stop words to reduce the BOW (bag
of words) size we also used lemmatization and stemming to be able to
reduce the number of words without taking off any valuable information.

\textbf{\emph{Lemmatization}} and different forms of
\textbf{\emph{Stemming}} were only used with \textbf{\emph{TF-IDF}}
while doing the Keywords extraction, otherwise we prefered keeping the
words as they were because we were doing a mapping with a pretrained
Stanford Word Embedding \cite{GloVe_web} and sometimes root forms were
not in the dictionnary of words of the pre-trained words-vectors.

For the Keywords extraction we used three methods : The first one is a
trivial approach that consists of searching in the pdf itself for
keywords and extract them, this method is the most accurate but we found
out that for 492 papers we had no keywords included. So we deciced to
use other two methods for keywords extraction, namely one based on
\textbf{\emph{TF-IDF}} and a second one based on \textbf{\emph{Rake}}.
\textbf{\emph{TF-IDF}} is one of the most popular term-weighting schemes
today, in information retrieval, tf--idf or TFIDF, short for term
frequency--inverse document frequency, is a numerical statistic that is
intended to reflect how important a word is to a document in a
collection or corpus. It is often used as a weighting factor in searches
of information retrieval, text mining, and user modeling. The tf--idf
value increases proportionally to the number of times a word appears in
the document and is offset by the number of documents in the corpus that
contain the word, which helps to adjust for the fact that some words
appear more frequently in general \cite{Tf-idf}. With TF-IDF we sorted
for each document the words with the highest score and then selected the
7 first words of each document.

    \textbf{\emph{Rake}}: Rake stands for Rapid Automatic Keyword
Extraction. It is an existing python implementation that uses the NLTK
toolkit for the score calculations.

\textbf{\emph{How the method works}} :

@@

Once we have all the keywords from the three methods, we tried to select
the most accurate keywords. So we created a method that selected the
intersection of TF-IDF keywords and the Rake keywords with the keywords
of the authors (if there are any keywords in the text), if the
intersection is empty and we have the keywords in the text we only take
the keywords of the text. If the paper does not contain any keywords, we
take the intersection of the TF-IDF and the Rake outputs. If again, the
intersection is empty then we only take the TF-IDF keywords because they
were closer to the text keywords than the Rake Keywords.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_21_0.png}\end{center}
        \caption{Trends of keywords over the years}
        \label{fig1}
    \end{figure}

    We wanted then to know what are the trends of words for each specific
year, so we selected distinct words for each year. In other words it
means that we selected for each year the words that were only present
for this specific year and never appeared in the others. We get the
figure below. As we can see from the graph \ref{fig2}, the most used
word in the papers, more than 10\% of 2015 papers used the word `peer',
more than 6\% of the papers used the words `building', `embodied' in
2016. In 2017 the most frequent keyword is `reponsive' and in 2018
`tools' is present with more than 6\%.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_23_0.png}\end{center}
        \caption{Distinct keywords over the years}
        \label{fig2}
    \end{figure}

@@

To cluster the documents we decided to use two different approaches. The
first one is based on Natural Langage Processing (NLP) and the second
one is rooted in network analysis. We fin that network clustering using
the references graph is most accurate.

@@

\hypertarget{baseline---tf-idf-combined-with-k-means}{%
\paragraph{Baseline - TF-IDF combined with
K-means}\label{baseline---tf-idf-combined-with-k-means}}

With the natural langage processing, we started with a naive method
which consisted of clustering the documents using only TF-IDF. Each
document has a vector representation with values for each word present
in the document depending on the frequency of the word in this document
regarding its frequency in all the other documents. So if a word is a
keyword to the document and is not a common word than its value will be
more important than the other words in the vector of the document. Once
we obtained our vector representation, we selected the max features
parameter as 2500, it means that we only considered the top
max\_features ordered by term frequency across the corpus. On this word
representation we run K-means.

\textbf{\emph{K-means}} clustering is a method used for clustering
analysis, especially in data mining and statistics. It aims to partition
a set of observations into a number of clusters (k), resulting in the
partitioning of the data into Voronoi cells. It can be considered as a
method of finding out which group a certain object really belongs to.

This method did not give good results as we can see it in the graph
below \ref{fig3}. All the vectors are together there is no clear
separation between the documents. So we decide to use other methods
using words embeddings combined with properties of TF-IDF.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_26_0.png}\end{center}
        \caption{Baseline model document clusters}
        \label{fig3}
    \end{figure}

@@

The second method was to use pre-trained words embeddings. We decided to
use \emph{GloVe} library pre-trained on Wikipedia dataset.

@@

We constructed a dictionnary from the words embeddings mapping each word
of a document to its vector representation, then we averaged all the
vectors of each document. Since each words has its own representation in
space, the average of all the words for each document will lead to a
vector representing a document.

Once we obtained a document vector representation, we applied K-means to
it. The results were slightly better, but we had only two clear
separated clusters. If we increase the number of clusters, then there is
no clear separation between clusters visible anymore. The plot in figure
\ref{fig4} reprensents the documents belonging to each cluster.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_28_0.png}\end{center}
        \caption{Cluster using Average of GloVe vectors combined with K-means }
        \label{fig4}
    \end{figure}

@@

With this method, we did not use the parameter max features of the
TF-IDF vector, we took all the vocabulary generated by all the documents
and we created a method that sorts the document frequency vector. We
selected the 500 first words and computed a weighted average of the
every word embedding vector with the its respectively TF-IDF score.

Once we get the weighted average vector, since we were using a words
embedding of 50 dimentions, we decided to use feature reduction
technique to reduce the number of dimensions and represent our vectors
in an other vector space where all the features are orthogonal to each
others.

To apply feature reduction to our vectors we used
\textbf{\emph{Principal Component Analysis (PCA)}} algorithm.

@@

Once we get our new vector representation, we applied K-means and we
obtained the best separation, with a clear separation of 4 clusters as
we can see it in figure \ref{fig5}.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_30_0.png}\end{center}
        \caption{Cluster using Weighted average of TF-IDF + GloVe  combined with K-means }
        \label{fig5}
    \end{figure}

@@

For the references graph we chose to construct 4 different graphs to be
able to extract as much information as possible. Firstly, to be able to
cluster the documents that are about the same subject and secondly to
gain more insights about the most influencial papers and trends of
citations.

@@

Our first graph is constructed using the references of each document. It
is constructed as follows: for each two documents, if there is an
intersection between the references of both documents, then there is a
link conneting the two documents. The weights of the edges are the
length of the intersection set. The intuition behind it is that the
higher the weight, the more the documents are likely to be about the
same subject, since they are citing the same documents. From this graph,
we create a subgraph of it. We filter the edges keeping only those who
have a weight strictly higher than 3. This means that, each node of the
subgraph, is related to at least one other node with 4 same documents
referenced.

    The resulting graph, plotted in figure \ref{intersection}, is an
undirected graph with 805 nodes and 6699 edges. On this graph we
extracted two main features:

\begin{itemize}
\tightlist
\item
  Strongly connected components to be able to detect clusters and
  related documents.
\item
  The in-degree in order to check if there are documents that are
  related to a lot of documents and if those documents are more general
  than others, and will not necessarly belong to one specific cluster.
\end{itemize}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_36_0.png}\end{center}
        \caption{Intersection between references graph}
        \label{intersection}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_57_0.png}\end{center}
        \caption{Global spread of community. Size of dot represent number of participants from each institution}
        \label{global_com}
    \end{figure}
